---
title: "Question_1"
author: "Ella Yang, Cecilia Jiang"
output:
  html_document: default
  pdf_document: default
---
```{r}
#Required Library
library(dplyr)
library(ggplot2)
```

```{r}
# Data Filtration
data <- background_clean
data <- data %>%
  mutate(
    pstat_courses_total = rowSums(select(., matches("PSTAT|pstat")), na.rm = TRUE)
  )
table(data$pstat_courses_total)
table(data$stat.prof)
```

```{r}
# Pie chart data preparation
prof_summary <- data %>%
  group_by(stat.prof) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(prof_summary, aes(x = "", y = percentage, fill = stat.prof)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 4) +
  labs(
    title = "Distribution of Students by Statistics Proficiency Level",
    x = NULL,
    y = NULL,
    fill = "Proficiency Level"
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

```{r}
#Boxplot
ggplot(data, aes(x = stat.prof, y = pstat_courses_total, fill = stat.prof)) +
  geom_boxplot() +
  labs(
    title = "Number of PSTAT Courses by Proficiency Level",
    x = "Proficiency Level",
    y = "Number of PSTAT Courses"
  ) +
  theme_minimal()

```

```{r}
#Average PSTAT Courses by Proficiency Level
data_summary <- data %>%
  group_by(stat.prof) %>%
  summarize(mean_courses = mean(pstat_courses_total, na.rm = TRUE),
            sd_courses = sd(pstat_courses_total, na.rm = TRUE))

ggplot(data_summary, aes(x = stat.prof, y = mean_courses, fill = stat.prof)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_courses - sd_courses, ymax = mean_courses + sd_courses), width = 0.2) +
  labs(
    title = "Average PSTAT Courses by Proficiency Level",
    x = "Proficiency Level",
    y = "Average Number of PSTAT Courses"
  ) +
  theme_minimal()
```


```{r}
data_summary <- data_summary %>%
  mutate(stat.prof = factor(stat.prof, levels = c("beg", "int", "adv")))

ggplot(data_summary, aes(x = stat.prof, y = mean_courses, fill = stat.prof)) +
  geom_col(width = 0.6) +
  geom_line(aes(group = 1), color = "darkblue", linewidth = 1) +
  geom_point(color = "darkblue", size = 3) +
  labs(
    title = "Average PSTAT Courses by Proficiency Level (Trend)",
    x = "Proficiency Level",
    y = "Average Number of PSTAT Courses"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

```


```{r}
# check numeric
data$stat.prof_num <- factor(data$stat.prof,
                         levels = c("beg", "int", "adv"),
                         labels = c("Beginner", "Intermediate", "Advanced"))
data$stat.prof_num <- as.numeric(data$stat.prof_num)
```

```{r}
# anova
model1 <- aov(pstat_courses_total ~ stat.prof_num, data = data)
summary(model1)
```
We found that the p-value (p = 0.00842) is less than the significance level $\alpha$ = 0.05, indicating a statistically significant difference in the number of PSTAT courses taken across different levels of statistics proficiency.

```{r}
# TukeyHSD
model2 <- aov(pstat_courses_total ~ stat.prof, data = data)
TukeyHSD(model2)
```
The Tukey HSD post-hoc test revealed significant differences among proficiency levels.

Advanced students had taken significantly more PSTAT courses than beginners (p = 0.0062 < $\alpha$ = 0.05), and intermediate students also took more courses than beginners (p = 0.033 < $\alpha$ = 0.05).

However, there was no significant difference between intermediate and advanced students (p = 0.59 > $\alpha$ = 0.05).
These results suggest that greater coursework exposure is associated with higher statistical proficiency, especially when comparing beginners to the other groups.

# Summary
In this analysis, we explored the relationship between students' statistics proficiency level and the number of PSTAT courses that they have taken.

Through descriptive statistics and visualization, we could conclude that students with higher proficiency levels generally completed more PSTAT courses. The one-way ANOVA confirmed that this difference was statistically significant (p = 0.008 < $\alpha$ = 0.05). Further, Tukey’s HSD post-hoc comparisons revealed that both intermediate and advanced students had taken significantly more PSTAT courses than beginners, while no significant difference was found between intermediate and advanced groups.

These findings indicate that greater exposure to PSTAT coursework is positively associated with higher levels of statistical proficiency, emphasizing the importance of continuous coursework in developing stronger data analysis skills.

